---
output: html_document
---
Lesson 4
========================================================

***

### Scatterplots and Perceived Audience Size
Notes: could also read .tsv with: **read.delim('pseudo_facebook.tsv')**
Could use ggplot for scatter as well:

  **ggplot(aes(x = age, y = friend_count), data = pf) +**
    **geom_point()**

***

### Scatterplots
Notes:

```{r Scatterplots}
library(ggplot2)
pf <- read.csv('pseudo_facebook.tsv', sep = '\t')

qplot(x = age, y = friend_count, data = pf)
qplot(age, friend_count, data = pf)
```

***

#### What are some things that you notice right away?
Response: Looking at the scatter plot you see stripes of vertical lines. i.e. people lied about age (100, 69 etc). Most people with lots of friends are in the 20s

***

### ggplot Syntax

```{r ggplot Syntax}
ggplot(aes(x= age, y = friend_count), data = pf) +
  geom_point() +
  xlim(13,90)

summary(pf$age)
```

***

### Overplotting
Notes: our graph has overplotting, so frak with alpha levels to render transparency. Done in geom point(). Also look at geom point() vs geom jitter().

```{r Overplotting}
ggplot(aes(x= age, y = friend_count), data = pf) +
  geom_jitter(alpha=1/20) + # takes 20 points for a circle to appear 100% opaque
  xlim(13,90)
```

#### What do you notice in the plot?
Response: Still large amount of age 69ers and most people are in their 20s. Slight decrease at age 40ish.Also friend counts are not as high as they use to be for people in their 20s, thanks to jitter vs point.

***

### Coord_trans()
Notes:

```{r Coord_trans()}
ggplot(aes(x= age, y = friend_count + 1), data = pf) +
  geom_jitter(alpha=1/20) + # takes 20 points for a circle to appear 100% opaque
  xlim(13,90) +
  coord_trans(y = "sqrt")
```

#### Look up the documentation for coord_trans() and add a layer to the plot that transforms friend_count using the square root function. Create your plot!

```{r}
ggplot(aes(x= age, y = friendships_initiated), data = pf) +
   geom_point(alpha=1/10, position = position_jitter(h=0)) + # takes 20 points for a circle to appear 100% opaque
   # xlim(13,90) 
   coord_trans(y = "sqrt")
```

#### What do you notice?

***

### Alpha and Jitter
Notes:

```{r Alpha and Jitter}

```

***

### Overplotting and Domain Knowledge
Notes:

***

### Conditional Means
Notes:

```{r Conditional Means}

```

Create your plot!

```{r Conditional Means Plot}

```

***

### Overlaying Summaries with Raw Data
Notes:

```{r Overlaying Summaries with Raw Data}

```

#### What are some of your observations of the plot?
Response:

***

### Moira: Histogram Summary and Scatterplot
See the Instructor Notes of this video to download Moira's paper on perceived audience size and to see the final plot.

Notes:

***

### Correlation
Notes:

```{r Correlation}

```

Look up the documentation for the cor.test function.

What's the correlation between age and friend count? Round to three decimal places.
Response:

***

### Correlation on Subsets
Notes:

```{r Correlation on Subsets}
#with(                 , cor.test(age, friend_count))
```

***

### Correlation Methods
Notes:

***

## Create Scatterplots
Notes:

```{r}

```

***

### Strong Correlations
Notes:

```{r Strong Correlations}

```

What's the correlation betwen the two variables? Include the top 5% of values for the variable in the calculation and round to 3 decimal places.

```{r Correlation Calcuation}

```

Response:

***

### Moira on Correlation
Notes:

***

### More Caution with Correlation
Notes:

```{r More Caution With Correlation}
# install.packages('alr3')
# library(alr3)
```

Create your plot!

```{r Temp vs Month}

```

***

### Noisy Scatterplots
a. Take a guess for the correlation coefficient for the scatterplot.

b. What is the actual correlation of the two variables?
(Round to the thousandths place)

```{r Noisy Scatterplots}

```

***

### Making Sense of Data
Notes:

```{r Making Sense of Data}

```

***

### A New Perspective

What do you notice?
Response:

Watch the solution video and check out the Instructor Notes!
Notes:

***

### Understanding Noise: Age to Age Months
Notes:

```{r Understanding Noise: Age to Age Months}

```

***

### Age with Months Means

```{r Age with Months Means}

```

Programming Assignment
```{r Programming Assignment}

```

***

### Noise in Conditional Means

```{r Noise in Conditional Means}

```

***

### Smoothing Conditional Means
Notes:

```{r Smoothing Conditional Means}

```

***

### Which Plot to Choose?
Notes:

***

### Analyzing Two Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

