Lesson 5
========================================================

### Moira Perceived Audience Size Colored by Age
Notes: still have horizontal stripes at 50, 100 etc (common guesses) bu there was no pattern in the data (adding color for age) so this looked to be a dead end.There was a lot of overplotting so its tough to see what more insight color can offer us.

***

```{r Create new DataFrame}
library(dplyr)
# chain functions together using %>%

pf.fc_by_age_gender <- pseudo_facebook %>%
  filter(!is.na(gender)) %>% #filters out NA for gender
  group_by(age, gender) %>%
  summarise(mean_friend_count = mean(friend_count), #summarise removes one layering of group
            median_friend_count = median(friend_count),
            n = n()) %>%
  ungroup() %>% #to remove other layering
  arrange(age)

head(pf.fc_by_age_gender)


```

### Third Qualitative Variable
Notes:

```{r Third Qualitative Variable}

ggplot(aes(x = gender, y = age),
       data = subset(pf.fc_by_age_gender, !is.na(gender))) + geom_boxplot() +
  stat_summary(fun.y = mean, geom = 'point', shape = 4)

ggplot(aes(x = gender, y = friend_count),
       data = subset(pf.fc_by_age_gender, !is.na(gender))) + 
  geom_line(aes(color = gender), stat = 'summary', fun.y = median)
```

***

### Plotting Conditional Summaries
Notes:

```{r Plotting Conditional Summaries}

ggplot(aes(x = age, y = median_friend_count), 
              data = subset(pf.fc_by_age_gender, !is.na(gender))) +   
  geom_line(aes(color = gender), stat = 'summary', fun.y = median)

```

***

<!-- ### Thinking in Ratios -->
<!-- Notes: -->

<!-- *** -->

### Wide and Long Format
Notes: Transpose data. E.g. columns of male/ female for each age. let's reshape it to one age for each row and then a male column and a female column so we can explore the friend count diff between male and female.

***

### Reshaping Data
Notes:

```{r}
install.packages('reshape2')
library(reshape2)

#dcast becauise it is D for Dataframe
pf.fc_by_age_gender.wide <- dcast(pf.fc_by_age_gender,
                                  age ~ gender,
                                value.var = 'median_friend_count')
# left side of ~ is from first DF that is what data you want to keep, chain with a +
# right side is the data you want recasted, in this case it is gender
# value.var allows us to set the data for gender from the first DF

head(pf.fc_by_age_gender.wide)

```


***

### Ratio Plot
Notes:

```{r Ratio Plot}

ggplot(aes(x = age, y = female / male),
       data = pf.fc_by_age_gender.wide) +
  geom_line() + 
  geom_hline(yintercept = 1, alpha = 0.3, linetype = 2)


```

***

### Third Quantitative Variable
Notes:

```{r Third Quantitative Variable}

```

***

### Cut a Variable
Notes:

```{r Cut a Variable}

```

***

### Plotting it All Together
Notes:

```{r Plotting it All Together}

```

***

### Plot the Grand Mean
Notes:

```{r Plot the Grand Mean}

```

***

### Friending Rate
Notes:

```{r Friending Rate}

```

***

### Friendships Initiated
Notes:

What is the median friend rate?

What is the maximum friend rate?

```{r Friendships Initiated}

```

***

### Bias-Variance Tradeoff Revisited
Notes:

```{r Bias-Variance Tradeoff Revisited}

ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure >= 1)) +
  geom_line(aes(color = year_joined.bucket),
            stat = 'summary',
            fun.y = mean)

ggplot(aes(x = 7 * round(tenure / 7), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 30 * round(tenure / 30), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 90 * round(tenure / 90), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

```

***

### Sean's NFL Fan Sentiment Study
Notes:

***

### Introducing the Yogurt Data Set
Notes:

***

### Histograms Revisited
Notes:

```{r Histograms Revisited}

```

***

### Number of Purchases
Notes:

```{r Number of Purchases}

```

***

### Prices over Time
Notes:

```{r Prices over Time}

```

***

### Sampling Observations
Notes:

***

### Looking at Samples of Households

```{r Looking at Sample of Households}

```

***

### The Limits of Cross Sectional Data
Notes:

***

### Many Variables
Notes:

***

### Scatterplot Matrix
Notes:

***

### Even More Variables
Notes:

***

### Heat Maps
Notes:

```{r}
nci <- read.table("nci.tsv")
colnames(nci) <- c(1:64)
```

```{r}
nci.long.samp <- melt(as.matrix(nci[1:200,]))
names(nci.long.samp) <- c("gene", "case", "value")
head(nci.long.samp)

ggplot(aes(y = gene, x = case, fill = value),
  data = nci.long.samp) +
  geom_tile() +
  scale_fill_gradientn(colours = colorRampPalette(c("blue", "red"))(100))
```


***

### Analyzing Three of More Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

